---
title: "QBA workshop - Exercises - Bayesian part"
author: Simon Dufour (Professor - Université de Montréal)
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography:
- book.bib
- packages.bib
biblio-style: apalike
link-citations: yes
description: Chapitre 2
output: html_document
---

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


# Generating prior distributions
One of the first thing that you will need for any Bayesian analysis is a way to generate and visualize a prior distribution corresponding to some scientific knowledge that you have about an unknown parameter. Generally, you will use information such as the mean and variance or the mode and the 2.5th (or 97.5th) percentile to find a corresponding distribution. You will use various distributions, for instance:  
  
**- Normal distribution:** defined by a mean (mu) and its standard deviation (SD) or variance (tau). In some software, OpenBUGS for instance, the inverse of the variance (1/tau) is used to specify a given normal distribution. **Notation: dNorm(mu, 1/tau)**  
**- Uniform distribution:** defined by a minimum (min) and maximum (max). **Notation: dUnif(min, max)**   
**- Beta distribution:** bounded between 0 and 1, beta distributions are defined by two shape parameters (a) and (b). **Notation: dBeta(a, b)**   
  
## Normal distribution
The `dnorm()` function can be used to generate a given Normal distribution and the `curve()` function can be used to visualize the generated distribution. These functions are already part of `R`, there is no need to upload a `R` package.  

```{r, message=FALSE, warning=FALSE, fig.cap="**Figure.** Density curve of a Normal distribution."}
curve(dnorm(x, mean=2.0, sd=0.5),                                    # I indicate mean and SD of the distribution
      from=-2, to=7,                                                 # I indicate limits for the plot
      main="Normal distribution with mean of 2.0 and SD of 0.5",     #Adding a title
      xlab = "Value", ylab = "Density")                              #Adding titles for axes

```
Note that a Normal distribution with mean of zero and a very large SD provides very little information. Such distribution would be referred to as a **uniform or flat distribution** (A.K.A.; a vague distribution).
```{r, message=FALSE, warning=FALSE, fig.cap="**Figure.** Density curve of a flat Normal distribution."}
curve(dnorm(x, mean=0.0, sd=10000000000),                                    
      from=-100, to=100,                                                 
      main="A flat Normal distribution",     
      xlab = "Value", ylab = "Density")                              

```
  
## Uniform distribution
In the same manner, we could visualize an uniform distribution using the `dunif()` function. In the following example, we assumed that any values between -5.0 and 5.0 are equally probable.
```{r, message=FALSE, warning=FALSE, fig.cap="**Figure.** Density curve of an Uniform distribution."}
curve(dunif(x, min=-5.0, max=5.0),                                    
      from=-10, to=10,                                                 
      main="Uniform distribution with -5.0 and 5.0 limits",    
      xlab = "Value", ylab = "Density")                              


```
  
## Beta distribution
Finally, Beta distributions are another type of distributions that will specifically be used for parameters that are proportions (i.e., bounded between 0.0 and 1.0). The `epi.betabuster()` function from the `epiR` library can be used to define a given prior distribution based on previous knowledge. When you use the `epi.betabuster()` function, it creates a new `R` object containing different elements. Among these, one element will be named *shape1* and another *shape2*. These correspond to the a and b shape parameters of the corresponding Beta distribution.   
  
For instance we may know that the most likely value for the sensitivity of a given test is 0.85 and that we are 95% certain that it is greater than 0.75. With these values, we will be able to find the a and b shape parameters of the corresponding Beta distribution.
```{r, message=FALSE, warning=FALSE, fig.cap="**Figure.** Density curve of a Beta distribution for a test sensitivity."}
library(epiR) 

# Sensitivity of a test as Mode=0.85, and we are 95% sure >0.75 
rval <- epi.betabuster(mode=0.85, conf=0.95, greaterthan=T, x=0.75)  # I create a new object named rval

rval$shape1                #View the a shape parameter in rval
rval$shape2                ##View the b shape parameter in rval  

#plot the prior distribution
curve(dbeta(x, shape1=rval$shape1, shape2=rval$shape2), from=0, to=1, 
      main="Prior for test's sensitivity", xlab = "Proportion", ylab = "Density")
```
  
Note that a dBeta(1.0, 1.0) is a uniform beta distribution.
```{r, message=FALSE, warning=FALSE, fig.cap="**Figure.** Density curve of a Beta(1.0, 1.0) distribution."}
#plot the prior distribution
curve(dbeta(x, shape1=1.0, shape2=1.0), from=0, to=1, 
      main="A Beta(1.0, 1.0) or flat distribution", xlab = "Proportion", ylab = "Density")
```

